\documentclass[11pt]{article}

% Change "review" to "final" to generate the final camera-ready version.
\usepackage[review]{acl}

% Standard ACL package includes.
\usepackage{times}
\usepackage{latexsym}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{microtype}
\usepackage{inconsolata}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{array}
\usepackage{amsmath}
\usepackage{placeins}

\title{A Layered Semantic DAG Compiler for Ancient Greek}
\author{Student Name \\
  New York University}

\begin{document}
\maketitle

\begin{abstract}
This paper presents a layered semantic compiler for Ancient Greek that follows
the original project proposal: raw text is processed through explicit
normalization, lexical, semantic-token, relation, and graph-construction layers,
and each layer remains inspectable and scoreable. The final representation is a
directed acyclic graph (DAG) whose nodes preserve token-level linguistic
information and whose edges use a compact semantic inventory:
\textsc{agent}, \textsc{theme}, \textsc{modifier}, \textsc{complement}, and
\textsc{coord}. I evaluate the same graph representation under three levels of
upstream support. In the fully raw-text setting, a supervised relation classifier
trained from official UD train splits reaches 85.89 lemma accuracy, 84.09 UPOS
accuracy, 86.25 proxy semantic-type accuracy, and 36.18 relation F1 on a
200-sentence held-out Perseus UD-proxy slice. In a dependency-free UD-token setting, where
gold UD tokens, lemmas, POS tags, and morphology are available but dependency
edges are hidden, the system reaches 62.41 F1 on a manually labeled historical
semantic gold set. In the UD-backed setting, where dependency edges are also
available, the semantic DAG compiler reaches 92.13 precision, 95.90 recall, and
93.98 F1 on the same manual historical gold set; a stricter leave-one-corpus-out
split gives 94.14 macro test F1. These results show that the semantic DAG layer
is strong when reliable syntactic structure is available, while fully automatic
relation induction from raw Greek remains the main bottleneck.
\end{abstract}

\section{Introduction}
Ancient Greek texts are often accessed through translations, commentaries, and
lexical notes, but many research questions require explicit structure over the
original language. A classic example is determining who did what to whom, under
which conditions, and with what modifiers. This is hard for a low-resource,
morphologically rich language because semantic interpretation depends on
surface form, lemma, morphology, word order, and syntactic attachment at the
same time.

The project proposal framed the system as a \emph{layered semantic compiler}
rather than as a black-box translator. The intended pipeline is:
\[
x \rightarrow z_{\text{phon}} \rightarrow z_{\text{morph}}
\rightarrow z_{\text{lex}} \rightarrow z_{\text{sem}}
\rightarrow E \rightarrow G,
\]
where \(x\) is the input text, intermediate \(z\)'s are structured layer outputs,
\(E\) is a labeled semantic edge set, and \(G\) is the final semantic DAG. This
paper keeps that framing. The research question is not only whether an
UD-annotated sentence can be converted into a graph, but how much semantic
graph quality changes as upstream linguistic support is removed.

The system therefore has three experimental settings. The raw-text setting uses
the full project pipeline and predicts relations without gold dependency edges.
The dependency-free UD-token setting provides gold tokens, lemmas, POS tags, and
morphology, but still withholds dependency edges. The UD-backed setting uses UD
dependency edges as an upstream syntactic analysis and compiles them into the
semantic DAG. All three settings share the same final graph representation and
evaluation protocol.

The paper makes four concrete contributions: a compact semantic DAG
representation for Ancient Greek, a transparent layered compiler that exposes
intermediate stages, a small manual semantic gold set over real historical Greek
sentences, and an empirical evaluation that separates raw-text relation
prediction from parser-backed semantic graph compilation.

\section{Related Work}
This work sits at the intersection of dependency annotation, classical
treebanking, semantic role labeling, and semantic graph representations.

Universal Dependencies provides a cross-linguistically consistent framework for
tokenization, part-of-speech tagging, morphological features, and dependency
relations \citep{nivre2016ud}. In this setting, UD matters because it supplies a
stable annotation layer from which semantic abstraction can begin. This makes it
possible to reorganize syntactic structure into semantic structure while preserving
the original token-level linguistic information.

The proposal also positioned this project against opaque subword and neural
pipelines. Byte-pair encoding was introduced for neural machine translation as a
way to handle rare and unknown words through subword units
\citep{sennrich2016bpe}. That solves a different problem from the one pursued
here: this project treats morphemes, lexemes, semantic token types, and semantic
relations as explicit intermediate objects. UDify \citep{kondratyuk2019udify}
shows that multilingual neural models can jointly predict morphology, lemmas,
and dependency syntax, but the present system instead emphasizes explicit
intermediate representations that can be inspected and scored layer by layer.

For Ancient Greek specifically, existing dependency treebanks make this kind of
work possible. The Ancient Greek and Latin Dependency Treebanks described by
\citet{bamman2011agldt} demonstrate the value of large, manually annotated
classical corpora for linguistic and philological research. The PROIEL project
\citep{haug2008proiel} extends the treebanking tradition to parallel and
historical Indo-European texts, including Ancient Greek material relevant to the
UD inputs used here. AGILe \citep{degraaf2022agile} is also relevant because it
shows that Ancient Greek lexical normalization benefits from
domain-sensitive, language-specific modeling. The treebank resources provide the
morphosyntactic basis for the system input used here, while AGILe motivates the
language-specific treatment of lexical normalization.

At the semantic level, the most relevant precedent is semantic role labeling.
PropBank-style annotation formalizes predicate-argument structure over events and
their participants \citep{palmer2005propbank}. The system adopts the
same general intuition that subject, object, and clausal dependencies can often
be reorganized into semantically meaningful role edges.

The approach is also inspired by graph-based semantic representations such as
Abstract Meaning Representation (AMR), which introduced a sentence-level
semantic representation for sembanking \citep{banarescu2013amr}. This paper
takes a similar graph-oriented perspective while using a compact inventory of
node types and role labels tailored to Ancient Greek treebank data. Related
graph parsers such as TUPA for UCCA \citep{hershcovich2017tupa} and graph-aware
AMR parsing models \citep{vasylenko2023amr} show that semantic parsing can be
treated as graph construction over structured semantic relations. The
experiments below therefore use labeled edge precision, recall, and F1 as the
main evaluation family.

What distinguishes the present system from the work above is its combination of
Ancient Greek UD input, a deliberately small semantic label inventory, explicit
graph validation, and evaluation directly on manually labeled historical Greek
sentences.

\section{Task Definition and Data}
\subsection{Task Definition}
The task is to map an Ancient Greek sentence into a labeled semantic DAG. Let
\(x = (w_1,\ldots,w_n)\) be an input token sequence. The output graph is
\[
G=(V,E), \qquad V=\{v_1,\ldots,v_n\},
\]
where each \(v_i\) stores linguistic information about token \(w_i\), and
\[
E \subseteq V \times V \times \mathcal{L}.
\]
The label inventory is
\[
\begin{aligned}
\mathcal{L} = \{&
\textsc{agent}, \textsc{theme}, \textsc{modifier},\\
&\textsc{complement}, \textsc{coord}\}.
\end{aligned}
\]
An edge \((v_i,v_j,\ell)\) means that token \(i\) is the semantic head of token
\(j\) under relation label \(\ell\).

The graph is intended to support downstream interpretability. A reader should be
able to inspect the output and recover the main event structure, participant
roles, modifier attachments, and clausal relationships without consulting the
original dependency labels. At the same time, the graph should remain
structurally valid: acyclic, internally consistent, and free of malformed edges.

\subsection{Evaluation Metrics}
For a predicted edge set \(E_p\) and gold edge set \(E_g\), exact labeled edge
precision, recall, and F1 are:
\[
P_E = \frac{|E_p \cap E_g|}{|E_p|}, \qquad
R_E = \frac{|E_p \cap E_g|}{|E_g|},
\]
\[
F1_E = \frac{2P_E R_E}{P_E + R_E}.
\]
The intersection is strict: source token, destination token, and label must all
match. This matches the proposal's plan to score relation prediction and full
DAG construction with labeled precision, recall, and F1.

The proposal also described a global layer-wise semantic score. In notation, a
weighted MUC-style objective can be written as:
\[
F_{\text{MUC}} =
\sum_{k \in K} \lambda_k F1_k,
\qquad
\sum_{k \in K}\lambda_k = 1,
\]
where \(K\) may include lexical normalization, semantic token assignment,
relation prediction, and DAG validity. In the experiments below, I report these
components separately rather than hiding them inside one weighted number. This
keeps the main bottleneck visible: raw relation prediction is substantially
harder than lexical and semantic-type prediction.

\subsection{Treebank Inputs}
The repository includes two public Ancient Greek UD test files used in the
experiments:

\begin{itemize}
  \item \texttt{data/real\_eval/}\\
  \texttt{grc\_perseus-ud-test.conllu}
  \item \texttt{data/real\_eval/}\\
  \texttt{grc\_proiel-ud-test.conllu}
\end{itemize}

When converted with the compiler, these files yield 2,353
sentences and 31,913 non-punctuation tokens. These treebanks provide the
morphosyntactic analyses from which the semantic DAGs are constructed.

For raw-pipeline training and proxy evaluation, the repository also supports
official PROIEL and Perseus train/dev/test CoNLL-U files under
\texttt{data/ud\_treebanks/}. These files are used as an external, ignored cache:
the relation classifier trains from the official train split, selects its
threshold on the dev split, and reports held-out performance on the test split.

\subsection{Manual Historical Semantic Gold}
The main semantic evaluation uses
\texttt{data/gold\_semantic/}\\
\texttt{historical\_semantic\_dags.jsonl}. This file contains
21 manually labeled historical Greek sentences selected from the PROIEL and
Perseus UD test materials. The examples total 236 tokens and 122 gold semantic
edges in the target inventory.

This manual gold set is intentionally content-semantic. It emphasizes event
participants, complements, coordination, and semantically relevant modifiers.
Articles, discourse particles, and other purely functional markers are not
systematically rewarded. That choice is important because a naive syntax-to-graph
mapping could otherwise inflate scores by reproducing formal syntactic detail
that does not contribute much to semantic interpretation.

The repository also evaluates a second view of the same manual gold set that
removes \textsc{modifier} from scoring. In that setting, the remaining core roles
(\textsc{agent}, \textsc{theme}, \textsc{complement}, and \textsc{coord}) account
for 94 gold edges. This core-role subset helps isolate predicate-argument and
clausal structure from broad modifier attachment.

\section{Methodology}
\subsection{Layered Raw Pipeline}
The raw pipeline follows the project proposal's modular design. A sentence is
processed by phoneme normalization, morpheme segmentation, lexeme normalization,
semantic token assignment, relation prediction, and DAG construction. Each stage
emits structured JSON consumed by the next stage. This matters for error
analysis: a relation error can be traced back to the relation classifier, to a
wrong semantic type, or to a lexical/morphological failure upstream.

The raw pipeline also uses transparent corpus-backed lexical hints. The
repository builds a fallback lexicon from local UD train files when they are
available. This gives the raw analyzer access to observed lemmas, UPOS tags, and
common morphological features without using gold dependency edges at inference
time. When exact forms are unavailable, suffix-based rules provide an
interpretable out-of-vocabulary fallback.

\subsection{Semantic Tokens and Nodes}
Regardless of the upstream setting, the final DAG uses the same node schema.
Each node stores the token surface form, lemma, part of speech, morphological
features, semantic type, semantic role, and confidence score. The semantic
typing scheme is intentionally compact. Nouns and proper nouns typically become
\emph{entity} nodes, verbs and auxiliaries become \emph{event} nodes, adjectives
become \emph{property} nodes, adverbs become \emph{manner} nodes, and conjunction
or linker-like items become \emph{connector} nodes. This inventory is designed
to support graph inspection, role attachment, and evaluation across real
historical sentences.

\subsection{Raw Relation Classifier}
The raw relation classifier is trained as a supervised token-pair classifier
from official UD train splits. For each candidate ordered pair \((i,j)\) within
a local window, the classifier predicts a label in
\[
\mathcal{L}' = \mathcal{L} \cup \{\textsc{none}\}.
\]
At inference time, for each dependent \(j\), the classifier keeps the highest
scoring non-\textsc{none} head-label pair if its score exceeds a tuned threshold:
\[
(\hat{i},\hat{\ell}) =
\arg\max_{i,\ell \in \mathcal{L}}
p(\ell \mid i,j,x),
\]
\[
p(\hat{\ell}\mid \hat{i},j,x) \geq \tau.
\]
The latest version adds contextual pair features: neighboring POS and lemma
features, whether a conjunction, preposition, determiner, or verb appears
between the candidate head and dependent, case/number/gender agreement, and
whether the candidate head is the nearest event or nominal around the dependent.
These features improve the raw relation classifier without changing the
UD-backed compiler.

\subsection{Dependency-to-Semantic Mapping}
The UD-backed compiler is a separate input branch. It does not run the raw
phoneme/morpheme/lexeme stages. Instead, it reads a CoNLL-U sentence, converts
UD token-level annotations into the semantic-token schema, and maps selected UD
dependency labels into the target semantic edge inventory. Table~\ref{tab:mapping}
summarizes the main mapping decisions.

\begin{table*}[!t]
\centering
\caption{Main mapping from UD dependency labels to semantic DAG labels.}
\label{tab:mapping}
\begin{tabular}{>{\raggedright\arraybackslash}p{0.28\textwidth} >{\raggedright\arraybackslash}p{0.18\textwidth} >{\raggedright\arraybackslash}p{0.45\textwidth}}
\toprule
UD labels & DAG label & Intuition \\
\midrule
\texttt{nsubj}, \texttt{csubj}, \texttt{obl:agent} & \textsc{agent} & canonical subjects and agentive obliques \\
\texttt{obj}, \texttt{iobj}, \texttt{nsubj:pass}, \texttt{csubj:pass}, \texttt{obl:arg} & \textsc{theme} & objects and patient-like arguments \\
\texttt{xcomp}, \texttt{ccomp}, \texttt{acl}, \texttt{advcl}, \texttt{advcl:cmp}, \texttt{obl} & \textsc{complement} & clausal and oblique complements \\
\texttt{amod}, \texttt{nmod}, \texttt{nummod}, \texttt{advmod}, \texttt{appos}, \texttt{det} & \textsc{modifier} & attributive, adverbial, or dependent modifiers \\
\texttt{conj} & \textsc{coord} & coordination between linked conjuncts \\
\bottomrule
\end{tabular}
\end{table*}

The mapping is intentionally conservative in its label inventory and broad in
coverage. Many syntactic distinctions are collapsed into a small number of
readable semantic edge types. Function-like determiners and discourse particles
are filtered when they would otherwise create semantically weak modifier edges.

Several UD relations are intentionally not treated as semantic edges. In the
corpus-scale coverage analysis, the largest unmapped relation types are
\texttt{case}, \texttt{cc}, \texttt{discourse}, \texttt{mark}, and \texttt{cop}.
These are mostly function markers or structural helpers rather than the central
content relations the graph is meant to expose. Excluding them makes the graph
more semantically readable and explains why content-dependency coverage is much
higher than overall dependency coverage.

\subsection{Graph Building and Validation}
After node and edge creation, both the raw and UD-backed branches call the same
DAG builder. The builder creates stable node identifiers such as \texttt{n0},
\texttt{n1}, and \texttt{n2}. Duplicate edges, self-loops, and malformed links
are removed before validation. Validation checks confirm that edge labels come
from the allowed inventory, endpoints are valid, and the final graph is acyclic.
The implementation also verifies predicate-argument well-formedness for required
structures and uses \texttt{networkx} for cycle detection.

These checks make the output usable as structured data, not just as a readable
annotation. High local edge accuracy is not enough if the graph as a whole is
invalid.

\section{Experiments}
\subsection{Experiment 1: Fully Raw UD-Proxy Evaluation}
The first experiment evaluates the raw pipeline. Sentences are reconstructed
from held-out UD test token forms and then passed through the project's raw
compiler. UD annotations are used only for evaluation: UD lemmas provide lexical
gold, UD UPOS tags provide POS gold, a deterministic UPOS-to-semantic-type map
provides proxy semantic-type gold, and selected UD dependencies provide proxy
semantic-relation gold. No UD dependency edges are given to the raw pipeline at
inference time.

The relation classifier is trained from official UD train splits, thresholded on
official dev splits, and evaluated on held-out test data. The reported
200-sentence run processes the test files in fixed order and exhausts the
sentence budget on \texttt{grc\_perseus-ud-test.conllu}; it should therefore be
read as a held-out Perseus UD-proxy slice rather than as a balanced full-treebank
score. The raw proxy relation metric also covers only the currently mapped
UD-derived labels \textsc{agent}, \textsc{theme}, \textsc{modifier}, and
\textsc{coord}; \textsc{complement} is evaluated in the manual semantic and
UD-backed settings below. This experiment is closest to the proposal's fully
automatic end-to-end setting, but its relation target is still a UD-derived proxy.

\subsection{Experiment 2: Manual Historical Semantic Gold}
The second experiment measures exact edge precision, recall, and F1 on the 21
manually labeled historical sentences. A predicted edge is counted as correct
only if its source token, destination token, and label all match the gold edge.
This metric is strict but appropriate because the output is a graph, not a bag
of independent labels. The experiment also reports DAG validity rate.

Two systems are compared on this task. The first is a dependency-free clean
baseline. It uses UD tokens, lemmas, POS tags, and morphological features, but
it ignores UD dependency links and predicts semantic edges from local word
order, POS, and case heuristics. The second is the UD-backed compiler, which
maps UD dependency edges into the semantic DAG label inventory. This comparison
separates the value of reliable lexical/morphological analysis from the value of
reliable dependency structure.

For both systems, two scores are reported. The \emph{full graph} score evaluates
all five labels. The \emph{core roles} score excludes \textsc{modifier} and
evaluates only \textsc{agent}, \textsc{theme}, \textsc{complement}, and
\textsc{coord}.

\subsection{Experiment 3: Corpus-Aware Historical Split}
A third experiment evaluates whether the manual-gold result depends on using all
manual examples for filter choices. The split evaluator performs
leave-one-corpus-out testing: it selects function-like modifier filters on one
corpus and tests on the other. This is a stricter check than a single pooled
manual-gold score because no held-out example is used during filter selection.

\subsection{Experiment 4: Corpus-Scale UD Conversion}
The final experiment measures how the UD-backed compiler behaves when applied to
the full combined Ancient Greek UD test data in the repository. Here the focus
is on coverage and robustness over the full converted corpus.

The relevant metrics are:
\begin{itemize}
  \item overall dependency coverage: the proportion of non-root,
  non-punctuation UD dependencies that receive a semantic edge;
  \item content dependency coverage: the same calculation, but excluding
  relations such as \texttt{case}, \texttt{cc}, \texttt{mark}, \texttt{cop},
  \texttt{discourse}, \texttt{aux}, \texttt{fixed}, and \texttt{flat:name};
  \item predicate-argument coverage: the proportion of subject, object, and
  argument-like dependencies that are converted into semantic edges;
  \item DAG validity rate: the proportion of converted sentences whose final
  graphs pass structural validation.
\end{itemize}

This experiment asks whether the system scales cleanly to thousands of real
treebank sentences while preserving the main semantic structure.

\subsection{Reproducibility}
The main experimental entry points in the repository are:
\begin{itemize}
  \item \texttt{scripts/train\_relation\_classifier.py} for the raw classifier;
  \item \texttt{scripts/evaluate\_historical\_gold.py} for manual semantic evaluation;
  \item \texttt{scripts/evaluate\_historical\_splits.py} for corpus-aware splits;
  \item \texttt{scripts/run\_ud\_dag.py} for corpus-scale conversion.
\end{itemize}
All quantitative results reported below were regenerated locally from those
scripts inside the cloned project workspace. The raw classifier result in
Table~\ref{tab:settings-results} comes from
\texttt{python scripts/train\_relation\_classifier.py --max-sentences 200
--output outputs/review\_relation\_classifier\_200.json}. That run selected
\(\tau=0.45\) on the dev split and reported 33.70 precision, 39.05 recall, and
36.18 F1 on the held-out test slice.

\section{Results}
\subsection{Overall Comparison}
Table~\ref{tab:settings-results} summarizes the three main experimental
settings. The important point is that these are not interchangeable scores:
they use the same final semantic DAG representation, but they differ in how much
upstream linguistic structure is available.

\begin{table*}[!t]
\centering
\caption{Main results across upstream-support settings.}
\label{tab:settings-results}
\begin{tabular}{>{\raggedright\arraybackslash}p{0.46\textwidth}rrrr}
\toprule
Setting & Precision & Recall & F1 & Support \\
\midrule
Raw text + classifier, Perseus UD-proxy slice & 33.70 & 39.05 & 36.18 & 1,091 \\
Dependency-free UD-token baseline, manual gold & 55.00 & 72.13 & 62.41 & 122 \\
UD-backed compiler, manual gold & 92.13 & 95.90 & 93.98 & 122 \\
UD-backed compiler, leave-one-corpus-out & 92.45 & 95.89 & 94.14 & 122 \\
\bottomrule
\end{tabular}
\end{table*}
The raw pipeline result is the most honest measure of the fully automatic
system. It reaches 36.18 relation F1 on the held-out Perseus UD-proxy slice. The same
run reports 85.89 lemma accuracy, 84.09 UPOS accuracy, and 86.25 proxy
semantic-type accuracy. This shows that the upstream lexical and semantic-token
layers are usable, while relation induction remains much harder.

The dependency-free UD-token baseline reaches 62.41 F1 on the manual historical
gold set. This is not a raw-text score, because gold tokenization, lemmas, POS
tags, and morphology are provided. However, it still withholds UD dependency
edges. The gap between 62.41 and 93.98 F1 therefore measures how much reliable
dependency structure contributes to semantic graph construction.

\subsection{Manual Historical Semantic Gold}
Table~\ref{tab:historical-results} gives the full and core-role comparison on
the manual historical gold set.

\begin{table*}[!t]
\centering
\caption{Manual historical semantic gold results on 21 Ancient Greek
sentences.}
\label{tab:historical-results}
\begin{tabular}{>{\raggedright\arraybackslash}p{0.44\textwidth}rrrr}
\toprule
System and setting & Precision & Recall & F1 & Gold edges \\
\midrule
Dependency-free baseline, full graph & 55.00 & 72.13 & 62.41 & 122 \\
UD-backed compiler, full graph & 92.13 & 95.90 & 93.98 & 122 \\
Dependency-free baseline, core roles only & 54.96 & 76.60 & 64.00 & 94 \\
UD-backed compiler, core roles only & 91.75 & 94.68 & 93.19 & 94 \\
\bottomrule
\end{tabular}
\end{table*}
The comparison is decisive. On the full graph, the UD-backed compiler improves
over the dependency-free baseline by 31.57 F1 points. On core roles, it improves
by 29.19 F1 points. This gap shows that the dependency layer is doing real work,
especially for long-distance event structure and clausal attachment.

\subsection{Per-Label Breakdown}
Table~\ref{tab:label-results} breaks down the UD-backed compiler's per-label
results on the same historical evaluation set.

\begin{table*}[!t]
\centering
\caption{Per-label results for the full compiler on the manual historical gold
set.}
\label{tab:label-results}
\begin{tabular}{lrrrr}
\toprule
Label & Precision & Recall & F1 & Gold edges \\
\midrule
AGENT & 80.95 & 100.00 & 89.47 & 17 \\
THEME & 93.33 & 90.32 & 91.80 & 31 \\
MODIFIER & 93.33 & 100.00 & 96.55 & 28 \\
COMPLEMENT & 94.59 & 94.59 & 94.59 & 37 \\
COORD & 100.00 & 100.00 & 100.00 & 9 \\
\bottomrule
\end{tabular}
\end{table*}
The strongest result is the compiler's performance on the semantic backbone of
the sentence. Core roles reach 93.19 F1, with particularly strong performance on
\textsc{complement} and perfect performance on \textsc{coord} in this sample.
The revised modifier filter also substantially improves the full graph:
\textsc{modifier} reaches 96.55 F1 rather than behaving as the dominant source
of over-generation.

Every evaluated historical sentence still yields a valid DAG. The validity rate
is 100\%, which means the structural layer of the system is robust even when the
semantic edge inventory is imperfect.

\subsection{Corpus-Aware Historical Split}
The leave-one-corpus-out experiment gives a macro test score of 94.14 F1. The
held-out Perseus fold scores 94.85 F1, and the held-out PROIEL fold scores 93.42
F1. This suggests that the manual-gold improvement is not only an artifact of
tuning filter choices on the same examples being evaluated.

\subsection{Corpus-Scale Conversion}
Table~\ref{tab:coverage-results} presents the corpus-scale conversion results on
the full combined Ancient Greek UD test material.

\begin{table*}[!t]
\centering
\caption{Corpus-scale conversion statistics on the combined Ancient Greek UD
test files in the repository.}
\label{tab:coverage-results}
\begin{tabular}{>{\raggedright\arraybackslash}p{0.36\textwidth}r>{\raggedright\arraybackslash}p{0.38\textwidth}}
\toprule
Metric & Value & Notes \\
\midrule
Sentences converted & 2,353 & Perseus + PROIEL test material \\
Tokens converted & 31,913 & non-punctuation tokens \\
Mapped dependencies & 23,458 & semantic edges produced \\
Unmapped dependencies & 6,102 & mostly function markers \\
Overall dependency coverage & 79.36\% & all non-root, non-punctuation deps \\
Content dependency coverage & 98.98\% & excluding function-marker relations \\
Predicate-argument coverage & 99.96\% & subject/object/argument-style deps \\
DAG validity rate & 100.00\% & all converted graphs valid \\
\bottomrule
\end{tabular}
\end{table*}
These coverage results provide a separate scale and robustness check for the
UD-backed adapter. They do not replace manual semantic evaluation, but they show
that the same adapter scales to thousands of real treebank sentences while
preserving nodes, retaining content dependencies, and producing structurally
valid DAGs.

The unmapped dependency counts support that interpretation. The largest unmapped
categories are \texttt{case} (2,322), \texttt{cc} (1,660), \texttt{discourse}
(788), \texttt{mark} (580), and \texttt{cop} (490). These are mostly structural
or functional relations rather than the main semantic relations the system is
trying to expose. Once such categories are excluded, content dependency coverage
rises to 98.98\%.
\FloatBarrier

\section{Discussion}
The most important result is the separation between representation quality and
raw prediction quality. The UD-backed compiler reaches 93.98 F1 on manual
historical semantic gold and 94.14 macro test F1 under leave-one-corpus-out
evaluation. This supports the claim that the proposed semantic DAG inventory is
a useful abstraction over real Ancient Greek syntactic structure.

The raw pipeline tells a different story. It reaches strong lexical and
semantic-type proxy scores, but only 36.18 relation F1. In other words, the
project has a working end-to-end pipeline, but relation induction from raw Greek
is still much harder than assigning lemmas, POS tags, or semantic token types.
This is also consistent with Ancient Greek word-order freedom: local token-pair
features help, but they do not fully recover the dependency structure needed for
semantic graph edges.

The dependency-free UD-token baseline sits between those two extremes. Its 62.41
F1 shows that reliable tokenization, lemma normalization, POS tags, and
morphology help substantially, even without dependency edges. At the same time,
the remaining gap to 93.98 F1 shows that explicit syntactic head information is
central to high-quality semantic graph construction.

This framing is important for the paper's claim. The system should not be
presented as a raw-text parser that achieves 93.98 F1. Instead, the contribution
is a transparent semantic-DAG framework with three evaluated operating points:
fully automatic raw text, dependency-free lexical/morphological support, and
parser-backed UD support. This matches the proposal's goal of a modular,
inspectable, scoreable pipeline.

\section{Conclusion and Future Work}
The paper presents a layered semantic DAG compiler for Ancient Greek. The system
follows the proposal's modular design: raw text can pass through explicit
phonemic, morphemic, lexical, semantic-token, relation, and DAG layers, while
UD-backed inputs can be used to test the semantic graph layer under stronger
upstream support. On 21 manually labeled historical sentences, the UD-backed
compiler reaches 93.98 F1 on the full graph and 93.19 F1 on core semantic roles,
compared with 62.41 and 64.00 F1 for a dependency-free UD-token baseline. In the
fully raw setting, the relation classifier reaches 36.18 F1 on held-out
UD-proxy relation evaluation, while lexical and semantic-type scores are above
84\%.

Future work should focus on the gap between raw relation prediction and
UD-backed semantic compilation. One direction is a structured head-selection
model that enforces one primary semantic head per dependent and optimizes graph
validity jointly with edge labels. Another is to expand the manual semantic gold
set so that raw, baseline, and UD-backed systems can be compared on the same
human-labeled target rather than relying on UD-proxy labels for the raw setting.
A third is to add richer semantic information such as valency classes, animacy,
or limited coreference while preserving the inspectable graph guarantees that
make the current system useful.

The result is a reproducible and modular Ancient Greek semantic compiler: the
raw pipeline is functional but still relation-limited, and the UD-backed
semantic DAG layer demonstrates that the proposed graph representation can
faithfully encode real historical Greek structure.

\section*{Limitations}
The main limitations are the small manual semantic gold set and the dependence
of the highest-scoring setting on upstream UD annotations. The raw UD-proxy
result is also limited: it is a 200-sentence Perseus test slice and scores only
selected UD-derived relation labels rather than the full manual semantic target.
The raw classifier itself is still simple: it predicts pairwise edges with local
context rather than performing global structured decoding. A more realistic raw
parser would likely need a two-stage architecture that first decides whether a
token should attach semantically, then selects the best head under graph-level
constraints.

\bibliography{references}

\end{document}
